---
title: "Assignment 1"
author: "Nicholas Knoblauch"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Homework Solutions

I am still in the process of writing up full homework solutions.  Hopefully this will get you started.


## Problem 11.3-4

Consider a hash table of size $m=1000$ and a corresponding hash function $h(k) = \lfloor m (kA \mod 1) \rfloor$ for $A = ( \sqrt{5}-1)/2$ Compute the locations to which the keys 61,62,63,64, and 65 are mapped

```{r engine='Rcpp'}
#include <Rcpp.h>
#include <math.h>





//[[Rcpp::export]]
int hash(const double k,const int m,const double A){
  return(std::floor((double)m*(fmod(k*A,1))));
}
```



```{r}
A = (sqrt(5)-1)/2
m = 1000L
hash(61,m,A)
hash(62,m,A)
hash(63,m,A)
hash(64,m,A)
hash(65,m,A)
```

## Problem 11.2-2

Demonstrate what happens when we insert the keys 5,28,19,15,20,33,12,17,10 into a hash table with collisions resolved by chaining.  Let the table have 9 slots, and let the hash function be $h(k)=k \mod 9$




```{r engine='Rcpp'}
#include <Rcpp.h>
#include <math.h>
#include <list>
// [[Rcpp::plugins(cpp11)]] 

//[[Rcpp::export]]
void simple_hash(Rcpp::IntegerVector keys, int num_slots=9){
  
  const int slots=num_slots;
  std::vector<std::list<int> > hash_table(slots,std::list<int>());
  
  for(auto key=keys.begin(); key!=keys.end(); key++){
    int k = *key;
    
    int hash_index = k % slots;
    
    Rcpp::Rcout<<"Key: "<<k<<" is being inserted at index: "<<hash_index<<std::endl;
    
    auto *hash_list = &hash_table[hash_index];

    hash_list->push_front(k);
    
    int post_size = hash_list->size();
    
    for(auto it=hash_list->begin(); it!=hash_list->end(); it++){
      Rcpp::Rcout<<*it<<"\t";
    }
    Rcpp::Rcout<<std::endl;
  }
}

```



```{r}

keys <- as.integer(c(5,
                     28,
                     19,
                     15,
                     20,
                     33,
                     12,
                     17,
                     10))
simple_hash(keys)
```

## Problem 10.1-2


```{r engine='Rcpp'}
#include <Rcpp.h>


//[[Rcpp::export]]
void two_stacks(const Rcpp::IntegerVector keys, const Rcpp::LogicalVector pushStackA, const Rcpp::LogicalVector popStackA){
  using namespace Rcpp;
  //The plan is to grow the stacks from opposite sides of the array
  
  if(keys.size()!=pushStackA.size()){
    Rcpp::stop("keys must be the same size as isStackA");
  }
  size_t n=keys.size();
  std::vector<int> stacks(n,0);
  
  int A_front=NA_INTEGER;

  int B_front=NA_INTEGER;


  //First we'll demonstrate pushing on to the stack
  for(int i=0; i<n; i++){
    if(pushStackA[i]==TRUE){
     if(Rcpp::IntegerVector::is_na(A_front)){
       A_front=n-1;
     }else{
       A_front--;
     }
     //If the tops of the stacks meet, we quit in protest
     if(A_front==B_front){
       Rcpp::stop("Out of room!");
     }
      Rcpp::Rcout<<"Pushing key["<<i<<"]"<<":("<<keys[i]<<") on stack A"<<std::endl;
      stacks[A_front]=keys[i];
    }else{
     if(Rcpp::IntegerVector::is_na(A_front)){
       B_front=0;
     }else{
       B_front++;
     }
     if(A_front==B_front){
       Rcpp::stop("Out of room!");
     }
      Rcpp::Rcout<<"Pushing key["<<i<<"]"<<":("<<keys[i]<<") on stack B"<<std::endl;
      stacks[B_front]=keys[i];
    }
  }
  
  int n_pop=popStackA.size();
  for(int i=0; i<n_pop; i++){
    if(popStackA[i]==TRUE){
      if(Rcpp::IntegerVector::is_na(A_front)){
        Rcpp::Rcout<<"Stack A is empty!"<<std::endl;
      }else{
        Rcpp::Rcout<<"Popping  element "<<stacks[A_front]<<" from stack A"<<std::endl;
        A_front++;
        if(A_front==n){
          A_front=NA_INTEGER;
        }
      }
    }else{
      if(Rcpp::IntegerVector::is_na(B_front)){
        Rcpp::Rcout<<"Stack B is empty!"<<std::endl;
      }else{
        Rcpp::Rcout<<"Popping  element "<<stacks[B_front]<<" from stack B"<<std::endl;
        B_front--;
        if(B_front==0){
          B_front=NA_INTEGER;
        }
      }
    }
  }
}

```


```{r}

keys <- 1:10

pushA <- sample(c(T,F),length(keys),replace=T)

popA <- sample(c(T,F),length(keys),replace=T)

two_stacks(keys,pushA,popA)


```


## Problem 2.1-3
At the start of the loop iteration $i$ we know that there is no element $A[j]=v$ for any $j<i$, (If this were true, we could not still be in the loop).  At the end of the loop iteration we know that $A[i]!=j$ (again, if this were true, we would have exited the loop).  After the loop we know we have either found the index of the element $v$, or we will be returning `NA` as these are the only two lines on which `ret_i` undergoes assignment. 


```{r engine='Rcpp'}

#include <Rcpp.h>


int linear_search(const Rcpp::IntegerVector A,const int v){
  
  bool isFound=false;
  size_t n=A.size();
  int ret_i=NA_INTEGER;
  for(size_t i=0; i<n; i++){
    if(A[i]==v){
      ret_i=i;
      isFound=true;
      break;
    }
  }
  return(ret_i);
}
```

## Problem 2.2-3

Let $p_i$ represent the probability that the $i$th element contains the target element $v$ ($p_i=p(A[i]=v)$.  If there are $n$ elements in $A$, then $p_i=1/n$.  Let $x_i=i$ represent the number of elements that are checked if the $i$th elemenent is the target element $v$.The smallest $x$ can be is 1, and the largest is $n$. Our expectation for $x$ is $$ \sum_{i=1}^n p_i x_i= \sum_{i=1}^N \frac{i}{n}=\frac{n(n+1)}{2n}=\frac{(n+1)}{2}$$ 





