---
title: "A Brief Note on Pointers"
author: "Nicholas Knoblauch"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


#An informal introduction.

As mentioned in the last discussion, computer memory can be thought of as a continuous array of `0`'s and `1`'s.  For example, my laptop has 8 gigabytes of memory.  A gigabyte is $10^9$ bytes, and a byte is 8 bits (a bit is one binary digit, a zero or a one).  This means that I can specify 64 billion binary digits in memory.  In theory, this could be represented as one (very large) integer between 0 and 64 billion.  While this is never done in practice, it is useful to consider the fact that  the "state" of a computer can be represented by the value of the memory at any given time.

## Pointer syntax.

In C++, as in any language, much of meaning is embedded in context.  In the example below, we assign the variable `x` the value of 2.  The assignment operation can be read as "x equals 2", but as we'll see, there's a little bit more going on here than what is implied by the equals sign.


```{r engine='Rcpp'}
#include <Rcpp.h>

//[[Rcpp::export]]
void pointerDemo(){
  int x=2;
  Rcpp::Rcout<<"The value of x is :"<<x<<std::endl;
  Rcpp::Rcout<<"The address of x is (in hexidecimal)"<<&x<<std::endl;
  int *xptr = &x;
  Rcpp::Rcout<<"The value of xptr is "<<xptr<<std::endl;
  Rcpp::Rcout<<"The address of xptr is "<<&xptr<<std::endl;
  Rcpp::Rcout<<"'de-referencing' xptr yields "<<*xptr<<std::endl;

  
  
  *xptr=3;
  Rcpp::Rcout<<std::endl<<std::endl;
  Rcpp::Rcout<<"The value of x is :"<<x<<std::endl;
  Rcpp::Rcout<<"The address of x is (in hexidecimal)"<<&x<<std::endl;
  Rcpp::Rcout<<"The value of xptr is "<<xptr<<std::endl;
  Rcpp::Rcout<<"The address of xptr is "<<&xptr<<std::endl;
  Rcpp::Rcout<<"'de-referencing' xptr yields "<<*xptr<<std::endl;
   
}

```



```{r}
pointerDemo()
```



## Pass by reference and pass by value

Working with relatively large vectors and matrices is very common in scientific code, this is especially true in R.
It's a good practice when writing code to break up code into discrete functions.  A function (usually) accepts some input and (usually) gives some output.  There are two ways to pass parameters to functions, by reference and by value.  R (in general) only allows pass by value.  

```{r engine='Rcpp'}

#include <Rcpp.h>


int pass_by_value(int x){
  x=x+1;
  return(x);
}


int pass_by_reference(int &x){
  x=x+1;
  return(x);
}



//[[Rcpp::export]]
void wrap_pass_by_value(){
  int x=4;
  Rcpp::Rcout<<"x is "<<x<<std::endl;
  int y=pass_by_value(x);
  Rcpp::Rcout<<"y is "<<y<<std::endl;
  Rcpp::Rcout<<"x is "<<x<<std::endl;
  
}

//[[Rcpp::export]]
void wrap_pass_by_reference(){
  int x=4;
  Rcpp::Rcout<<"x is "<<x<<std::endl;
  int y=pass_by_reference(x);
  Rcpp::Rcout<<"y is "<<y<<std::endl;
  Rcpp::Rcout<<"x is "<<x<<std::endl;
  
}


```


```{r}

wrap_pass_by_reference()
wrap_pass_by_value()


```



